---
title: "DDA"
output: html_document
date: "2023-03-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Any necessary libraries

```{r}

```

```{r}
df <- read.csv("train.csv")
```


#2. Data preparation an cleaning

The following points will be checked to assess the data quality:

- What type of data are we working with? Are data types correct? Do the values make sense?

FIrstly, we can eyeball the data to gain a general idea of what the variables we're working with look like:

```{r}
str(df)
```
At first sight, there are some variables we do not really need to answer our research question, so we can proceed to delete them:

```{r}
df <- df[, -c(1)]
df <- df[, -c(3)]
df <- df[, -c(5)]

#Note if any of you consider we can deleter more columns, add them here
```


- Are there erratic values? Are there more categories than expected for categorical variables?

- Are there any missing values? How am I going to deal with them?
