---
title: "EDA_data"
author: "Tanvi Mathur"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(corrplot)
library(GGally)
library(factoextra)
library(cluster)

```


#3. EDA

The following questions will have to be answered:

- How is each variable distributed?
- What is the correlation between variables? (Correlation matrix)
- We have to apply an unsupervised method (clustering perhaps?)

```{r}
clean_model <- read.csv("df_model.csv")
```

```{r}
summary(clean_model)
```

```{r}
# Check number of unique values for variable "Customer_ID"
length(unique(clean_model$Customer_ID))
```
There are around 103,010 values which are not unique.


# Variable Distribution

```{r}
# Bar chart for variable "Occupation"
ggplot(clean_model)+ geom_bar(aes(x=Occupation), colour="black", fill="light blue") +
  labs(title = "Occupation") +
  scale_y_continuous(name="Frequency") +
  theme_light()
```
the proportion of each occupation is roughly equal, then you can assume that the data is balanced.

```{r}
# Histogram for variable "Annual_Income"
ggplot(clean_model, aes(x=Annual_Income)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Annual Income") +
  theme_light()

```
The data is right skewed, that's the number of people decreases with higher annual incomes.
```{r}
# Histogram for variable "Monthly_Inhand_Salary"
ggplot(clean_model, aes(x=Monthly_Inhand_Salary)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Monthly In-hand salary") +
  theme_light()

```
The data is right skewed, that's the number of people decreases with higher monthly incomes.
```{r}
# Histogram for variable "Num_of_Loan"
ggplot(clean_model)+ geom_bar(aes(x=Num_of_Loan), colour="black", fill="light blue") +
  labs(title = "Number of Loans") +
  scale_y_continuous(name="Frequency") +
  theme_light()
```
There are more people with number of loans ranging from 2 to 4.

```{r}
# Histogram for variable "Delay_from_due_date"
ggplot(clean_model, aes(x=Delay_from_due_date)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Delay from due date") +
  ggtitle("Delay from due date") +
  theme_light()

```


```{r}
# Histogram for variable "Num_of_Delayed_Payment"
ggplot(clean_model, aes(x=Num_of_Delayed_Payment)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Number of delayed payments") +
  theme_light()

```


```{r}
# Histogram for variable "Outstanding_Debt"
ggplot(clean_model, aes(x=Outstanding_Debt)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Outstanding debt") +
  theme_light()

```

```{r}
# Histogram for variable "Credit_History_Age"
ggplot(clean_model, aes(x=Credit_History_Age)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Credit History age") +
  theme_light()

```


```{r}
# Histogram for variable "Payment_of_Min_Amount"
ggplot(clean_model)+ geom_bar(aes(x=Payment_of_Min_Amount), colour="black", fill="light blue") +
  labs(title = "Payment of min amount") +
  scale_y_continuous(name="Frequency") +
  theme_light()
```


```{r}
# Histogram for variable "Total_EMI_per_month"
ggplot(clean_model, aes(x=Total_EMI_per_month)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Total EMI per month") +
  theme_light()

```


```{r}
# Histogram for variable "Monthly_Balance"
ggplot(clean_model, aes(x=Monthly_Balance)) + 
  geom_histogram(aes(y=..density..),      
                 #binwidth=1,
                 colour="black", fill="light blue") +
  geom_density(alpha=.2, fill="blue") +   
  xlab("Monthly_Balance") +
  theme_light()

```

```{r}
# Histogram for Payment_Behaviour
ggplot(clean_model)+ geom_bar(aes(x=Payment_Behaviour), colour="black", fill="light blue") +
  labs(title = "Payment behaviour") +
  scale_y_continuous(name="Frequency") +
  theme_light()
```

#Correlation among numerical variables 

```{r}
corr_model <- clean_model[,!names(clean_model) %in% c("X", "ID","Customer_ID","Occupation", "Num_of_Loan", "Payment_of_Min_Amount", "Payment_Behaviour")]

# Calculate the correlation matrix
cor_matrix <- cor(corr_model)

# Create a correlation plot
corrplot(cor_matrix, method = "color",type = "upper",
        tl.col = "black",  tl.srt = 45, addCoef.col = "black")
```
outstanding correlations:
- Annual income and MOnthly inhand salary are prefectly correlated, so one of them can be eliminated prior to modeling
- Monthly balance and annual income
- Outstanding debt and delay from due date (as expected)
- Outstanding debt and credit history age seem to be negatively correlated, thus newer accounts seem to have more debt

# What further knowledge can we imply from the data?
 - Is there any occupation that is prone to delay in payment
```{r}
ggplot(clean_model, aes(x = Occupation, y = Num_of_Delayed_Payment)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Occupations with Delayed Payments",
       x = "Occupation",
       y = "Number of Delayed Payments")
```

 - Is there any occupation that has more loans than other

```{r}
ggplot(clean_model, aes(x = Occupation, y = Num_of_Loan)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Occupations with Delayed Payments",
       x = "Occupation",
       y = "Number of Delayed Payments")
```


# Unsupervised learning - cluster analysis

```{r}
# 2. cluster analysis
### 2.1 hierarchical clustering - complete linkage
###   note: exclude the Region attribute

#We rake a sample of the first 10k rows, as the dataset is too big and it would take much more time
dummy_df <- clean_model[1:10000,]

#we clean the first columns as they have no importance now (ID, customer ID, "X", and occupation)
dummy_df <- dummy_df[, -c(1,2,3,4)]


dist_df <- dist(dummy_df, method = 'euclidian')

#elbow method to check the optimal number of clusters
fviz_nbclust(dummy_df, kmeans, method = "wss")
```


```{r}
set.seed(2020)

k_df1 <- kmeans(dummy_df, 3, nstart = 25)
#25 different random starts to avoid missplacements and obtain a more robust solution

sil <- silhouette(k_df1$cluster, dist_df)
fviz_silhouette(sil)
```
As seen, there are no negative values for any cluster, which indicates that theres no missclassifications. moreover, the clusters are all above the avg coefficient line, which indicates that 3 might be the optimal number of clusters.
